---
title: "[Forecasting Notes](https://www.otexts.org/fpp/1)"
author: "Francis Tan"
date: "2016-08-31"
---

```{r include=FALSE}
require(fpp)
# Load the data
```

Chapter 1: Getting Started
==========================

## Types of Quantitative Forecasts
- Cross-sectional Data
	- Given a set of parameters, try to _predict_ an outcome based on data. For
example, predict the house price based on number of bedrooms, bathrooms, etc.
- Time series Data
	- Forecast future outcome based on historical data

## Basic Steps of Forecasting
1. Problem Definition
2. Gathering Information
3. Exploratory Analysis
4. Choosing and Fitting Models
5. Using and Evaluating Model


Chapter 2: Forecaster's Toolbox
===============================

## Graphs
First thing to do for any forecasting exercise is to plot the data to look for
patterns or any abnormalities.

### Time Plots
aka Line graphs.

#### Example 1
```{r}
data(melsyd)
plot(melsyd[,"Economy.Class"], 
  main="Economy class passengers: Melbourne-Sydney",
  xlab="Year",ylab="Thousands")
```
##### Notes
- Missing data in 1989 -- industrial dispute
- Dip in 1992 -- trial which replaced some economy class seats with business
class
- Large increase in 1991
- etc
